\LARGE \bf Energy-Aware Dynamic Mission Planning Algorithm for UAVs

### 2020-11-06 icra-2021.tex
finally implemented all the Ulrik's comments

    booktitle={2020 4th IEEE International Conference on Robotic Computing (IRC)},
    pages={to appear},

@@

    booktitle={2020 4th IEEE International Conference on Robotic Computing (IRC)},
    pages={to appear},

@@

\begin{figure*}[t]
  \centering
  \input{\figpath/icra-2021_overview2.tikz}
  \caption{The mission consists of circles and lines covering a polygon. The UAV heads to $\varphi_0$ initially, using the desired velocity vector $\dot{\mathbf{p}}_d$. It later heads to $\varphi_0^-$ reducing the radius $c_{0,1}$ to satisfy the battery constraints (indeed the energy coefficients $\mathbf{q}_{k_2}$ evolution is higher then $\mathbf{q}_{k_3}$). The UAV then converges to $\varphi_1$ in stage $\mathcal{M}_1$, $\varphi_2$ in stage $\mathcal{M}_2$, and so on (the circle $\varphi_2$ is not visible in the figure; it connects $\varphi_1$ and $\varphi_3$).}
  \label{fig:overview}
\end{figure*}

@@

Planning a mission for unmanned aerial vehicles (UAVs) operating outdoors is a challenging task. Scenarios such as precision agriculture, search and rescue, and surveillance require advanced levels of autonomy but have limited energy budgets. A typical example would be a UAV used to inform its ground-based actors of patterns detected while flying. Currently, UAVs flying outdoors are often semi-autonomous, in the sense that the mission is static and usually defined using mission planning software~\cite{daponte2019review}. Such a state of practice has prompted us to propose an \emph{energy-aware dynamic mission planning algorithm} for UAVs. The algorithm combines and generalizes some of the past body of knowledge on mobile robot planning problems, and addresses the increasing \emph{computational demands} and their relation to energy consumption, path, and autonomy.

Planning algorithms for mobile robots are a related to topics such as trajectory generation and path planning. Generally, these algorithms select an energy-optimized trajectory~\cite{mei2004energy}, e.g., by maximizing the operational time~\cite{wahab2015energy}, but in practice apply to few robots~\cite{kim2005energy}, and focus on optimizing motion control for these robots~\cite{kim2008minimum}, despite compelling evidence for the energy consumption also being influenced by on-board computations and not only bare motion~\cite{mei2005case}. For UAVs specifically, rotorcrafts have gained interest in terms of algorithms for energy-optimized trajectory generation~\cite{morbidi2016minimum,kreciglowa2017energy}. Mission planning algorithms, which includes a broader concept of a mission being a set of tasks along with a motion plan, all focus on the trajectory~\cite{mei2005case,mei2006deployment} and apply to few robots~\cite{sadrpour2013mission,sadrpour2013experimental}. Given the availability of powerful GPU-equipped mobile hardware, the use of on-board computations is expected to increase in the near future.
% todo cite
% todo add paragraph telling the algo (title)

The proposed algorithm alters the energy consumption dynamically by means of mission-specific parameters: the Quality of Service (QoS) of the on-board computations, and the trajectory-explicit equations (TEEs) adjustments. In the remainder of the paper, we adopt the following notation. We refer to the values of mission-specific QoS and TEEs parameters as computations and adjustments, we refer to the constraints sets that delimit such computations and adjustments as QoS and TEEs sets, and we refer to the current trajectory as TEE. Our goal is to complete the mission with highest possible computations and adjustments as the UAV flies and its batteries drain. The algorithm both optimizes computations in the form of robot operating system (ROS) nodes, and it optimizes adjustments--a way to alter the trajectory--guiding the UAV using a vector field~\cite{de2017guidance} that converges smoothly to such trajectory. It relies on the assumptions of the mission being \emph{periodic} and \emph{uncertain}. Periodicity is often observed, e.g., when the UAV is flying in repetitive patterns for a monitoring or search and rescue mission. Uncertainty accounts for the environmental interference, e.g., a fixed-wing UAV drifting due to windy weather. The algorithm addresses the periodicity by modeling the energy with Fourier analysis: the mission is periodic, so we expect the energy to also approximately evolve periodically. The uncertainty is addressed with a state estimator. The algorithm outputs the control input (adjustments and computations) using output model predictive control (MPC).

@@

We assume that the mission consists of different stages. At each stage the robot must follow a path and do some tasks.

@@

The algorithm uses the set from Equation~\text{eq:area} to select the optimal adjustments $\mathbf{c}_i^0$ s.t. $\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$ has the highest instantaneous energy consumption under the battery constraints. It guides the UAV to the new position $\mathbf{p}_{k+1}$ computing the vector field of the function $\varPhi:=\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$. It uses the vector field $\nabla\varPhi:=(\partial\varPhi/\partial x,\partial\varPhi/\partial y)$ to derive the direction to follow in the form of velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $E$ specifies the tracking direction, and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference, such as wind ($w\in\mathbb{R}$ in Figure~\ref{fig:overview}).

@@

Suppose there is no state and output uncertainty (the contributions $\mathbf{w}_k\in\mathbb{R}^m,v_k\in\mathbb{R}$ are zero). The non-linear model in Equation~(\ref{eq:fourier}) can be expressed using an equivalent linear discrete time-invariant state-space model
\begin{equation}\label{eq:state-perf}\begin{split}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+\mathbf{w}_k,\\
  y_k&=C\mathbf{q}_k+v_k,
\end{split}\end{equation}
where $y_k\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption. The state $\mathbf{q}$ mimics the original Fourier series coefficients
\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}_k&=\left[\begin{array}{cccccc}
    \alpha_0 & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{cccc}
    1&    &       &  \\
     & A_1&       &  \\
     &    & \ddots&  \\
     &    &       & A_r 
  \end{array}\right],\,A_i=\begin{bmatrix}0 & i/\xi \\ -i^2/\xi^2 & 0\end{bmatrix},\\
  C&=\left[\begin{array}{cccccc}
    1 & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}_k\in\mathbb{R}^m$ given $m=2r+1$, $A\in\mathbb{R}^{m\times m}$ is the state transmission matrix, and $C\in\mathbb{R}^m$ is the output matrix. In matrix $A$, the top left entry is one, the diagonal entries are defined by $A_1,\dots,A_r$, and zero the remaining entries are zero.

@@

Note from Equation~(\ref{eq:state-control}) that the control input $\mathbf{u}_k^a$ differs from the nominal control $\mathbf{u}_k$ in Equation~(\ref{eq:state-perf}). The first is an output of the algorithm. The second is a function that maps two consecutive control inputs to the difference in instantaneous energy consumption $\mathbf{u}_k:\mathbb{R}^{2\rho}\times\mathbb{Z}_{\geq 0}^{2\sigma}\rightarrow\mathbb{R}^{\rho+1}$.

@@

Specifically, the mission $\mathcal{M}$ contains a set of ordered lists with tasks (recall Definition~\ref{def:mission}). These tasks are implemented by software components (such as ROS nodes) $\Psi(\mathbf{s}_i):=\langle\psi_1(s_{i,1}),\dots,\psi_\sigma(s_{i,\sigma})\rangle$. They input the desired and output the actual computations
\begin{equation}\label{eq:qos-def}
  \mathbf{s}_i:=\left\{\langle s_{i,1},\dots ,s_{i,\sigma}\rangle\mid \psi_j(s_{i,j})\in\mathbb{S}_{i}\,\forall j\in[\sigma]^+\right\},
\end{equation}
where $s_{i,j}:\mathbb{Z}_{\geq 0}\rightarrow\mathbb{Z}_{\geq 0}$ returns the $j$-th desired computation at stage $i$, $\mathbf{s}_i\in\mathbb{S}_i\subseteq\mathbb{Z}_{\geq 0}^\sigma$ the union of all QoS sets given $\mathbb{S}_i:=\bigcup_{j\in[\sigma]^+}{\mathbb{S}_{i,j}}$.

For instance, if the task $\psi_1$ is a CNN object detector, $s_{1,1}$ corresponds to the computation frames-per-second (fps) rate changing the detection frequency. The algorithm output the control input which contains the desired fps rate $s_{1,1}\in\mathbb{S}_{1,1}$. The task $\psi_1(s_{1,1})$ output the actual fps rate $s_{k,1}$ which might differ from the desired one (e.g., the CNN object detector might not be able to reach a high fps rate with the current computational resources).

Let us further define $g:\mathbb{Z}_{\geq 0}^{\sigma}\rightarrow\mathbb{R}_{\geq 0}$ as the actual instantaneous computational energy consumption value obtained using \powprof{}
\begin{equation}\label{eq:energy-comp}\begin{split}
  y_k^s:=g\left(\Psi(\mathbf{s}_i)\right)=g\left(\mathbf{s}_{k}\right).\\
\end{split}\end{equation}
Moreover, $g(\{\emptyset\})=0$.

@@

Given $l$ stages $\mathcal{M}_i$ (TEE $\varphi_i$, tasks $\Psi$, computations $\mathbf{s}_i\in\mathbb{S}_i$, and adaptations $\mathbf{c}_i\in\mathbb{C}_i$) for all $i\in[l]$, the main purpose of the algorithm is to output a control input sequence $\mathbf{u}^a:=\{\mathbf{u}_0^a,\mathbf{u}_1^a,\dots\}$ in a valid mission.

@@

We stated earlier the output $y_k$--the instantaneous energy consumption--evolves in $\mathbb{R}_{\geq 0}$. This is generally untrue. Physical UAVs are bounded by strict battery limitations.

>>>>

    booktitle={2020 Fourth IEEE International Conference on Robotic Computing (IRC)},
    pages={135--142},

@@

    booktitle={2020 Fourth IEEE International Conference on Robotic Computing (IRC)},
    pages={294--297},

@@

\begin{figure*}[t]
  \centering
  \input{\figpath/icra-2021_overview2.tikz}
  \caption{The mission consists of circles and lines covering a polygon. The UAV heads to $\varphi_0$ initially, using the desired velocity vector $\dot{\mathbf{p}}_d$. It later heads to $\varphi_0^-$ reducing the radius $c_{0,1}$ to satisfy the battery constraints. The UAV then converges to $\varphi_1$ in stage $\mathcal{M}_1$, $\varphi_2$ in stage $\mathcal{M}_2$, and so on (the circle $\varphi_2$ is not visible in the figure; it connects $\varphi_1$ and $\varphi_3$).}
  \label{fig:overview}
\end{figure*}

@@

Planning a dynamic mission for unmanned aerial vehicles (UAVs) operating outdoors is a challenging task. Scenarios such as precision agriculture, search and rescue, and surveillance require advanced levels of autonomy but have limited energy budgets. A typical example of a mission with these limitations is a UAV flying a trajectory and performing some on-board computations. For instance, the UAV might detect some patterns and inform other, ground-based, actors. Planning a dynamic mission for such UAV involves finding optimal tradeoffs between the trajectory and computations. However, UAVs flying outdoors are often semi-autonomous, in the sense that the mission is static and usually defined using mission planning software~\cite{daponte2019review}. Such a state of practice has prompted us to propose an \emph{energy-aware dynamic mission planning algorithm} for UAVs. The algorithm combines and generalizes some of the past body of knowledge on mobile robot planning problems, and addresses the increasing \emph{computational demands} and their relation to energy consumption, path, and autonomy.

Planning algorithms for mobile robots are related to topics such as trajectory generation and path planning. Generally, these algorithms select an energy-optimized trajectory~\cite{mei2004energy}, e.g., by maximizing the operational time~\cite{wahab2015energy}, but in practice apply to a small number of robots~\cite{kim2005energy}, and focus on optimizing motion control for these robots~\cite{kim2008minimum}, despite compelling evidence for the energy consumption also being influenced by computations and not only bare motion~\cite{mei2005case}. For UAVs specifically, rotorcrafts have gained interest in terms of algorithms for energy-optimized trajectory generation~\cite{morbidi2016minimum,kreciglowa2017energy}. Mission planning algorithms, which include a broader concept of a mission being a set of tasks along with a motion plan, all focus on the trajectory~\cite{mei2005case,mei2006deployment} and apply to a small number of robots~\cite{sadrpour2013mission,sadrpour2013experimental}. Given the availability of powerful GPU-equipped mobile hardware, the use of computations is expected to increase in the near future.

The user specifies an initial mission plan--the trajectory and the tasks--along with an energy budget and the following mission specification parameters. The Quality of Service (QoS) computations parameters are relative to the tasks, and trajectory-explicit equations (TEEs) adjustments parameters to the path. The proposed algorithm replans the initial mission plan dynamically as the energy budget changes due to atmospheric interferences. In the remainder of the paper, we adopt the following notation. We refer to the values of mission-specific QoS and TEEs parameters as computations and adjustments, we refer to the constraints sets that delimit such computations and adjustments as QoS and TEEs sets, and we refer to the current trajectory as TEE. Our goal is to complete the mission with the highest possible computations and adjustments as the UAV flies and its batteries drain. The algorithm both optimizes computations in the form of robot operating system (ROS) nodes, and it optimizes adjustments--a way to alter the trajectory--guiding the UAV using a vector field~\cite{de2017guidance} that converges smoothly to such trajectory. It relies on the assumptions of the mission being \emph{periodic} and \emph{uncertain}. Periodicity is often observed, e.g., when the UAV is flying in repetitive patterns for monitoring or search and rescue scenarios. Uncertainty accounts for the atmospheric interference, e.g., a fixed-wing UAV drifting due to windy weather. The algorithm addresses the periodicity by modeling the energy with Fourier analysis: the mission is periodic, so we expect the energy to also approximately evolve periodically. The uncertainty is addressed with a state estimator. The algorithm outputs the control input (adjustments and computations) using output model predictive control (MPC).

@@

We assume that the initial mission plan consists of different stages. At each stage the robot must follow a path and do some tasks.

@@

The algorithm uses the set from Equation~(\ref{eq:area}) to select the optimal adjustments $\mathbf{c}_i^0$ s.t. $\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$ has the highest instantaneous energy consumption (while still respecting the energy budget). It guides the UAV to the new position $\mathbf{p}_{k+1}$ using the function $\varPhi:=\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$, computing its vector field $\nabla\varPhi:=(\partial\varPhi/\partial x,\partial\varPhi/\partial y)$, and deriving the direction to follow in the form of velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $E$ specifies the tracking direction, and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference, such as wind ($w\in\mathbb{R}$ in Figure~\ref{fig:overview}).

@@

\begin{lem}[Fourier series, harmonic oscillator equivalence]Suppose there is no state and output uncertainty (the contributions $\mathbf{w}_k\in\mathbb{R}^m,v_k\in\mathbb{R}$ are zero). The non-linear model in Equation~(\ref{eq:fourier}) can be expressed using an equivalent linear discrete time-invariant state-space model
\begin{equation}\label{eq:state-perf}\begin{split}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+\mathbf{w}_k,\\
  y_k&=C\mathbf{q}_k+v_k,
\end{split}\end{equation}
where $y_k\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption. The state $\mathbf{q}$ are the energy coefficients
\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}_k&=\left[\begin{array}{cccccc}
    \alpha_0 & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{cccc}
    1&    &       &  \\
     & A_1&       &  \\
     &    & \ddots&  \\
     &    &       & A_r 
  \end{array}\right],\,A_i=\begin{bmatrix}0 & i/\xi \\ -i^2/\xi^2 & 0\end{bmatrix},\\
  C&=\left[\begin{array}{cccccc}
    1 & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}_k\in\mathbb{R}^m$ given $m=2r+1$, $A\in\mathbb{R}^{m\times m}$ is the state transmission matrix, and $C\in\mathbb{R}^m$ is the output matrix. In matrix $A$, the top left entry is one, the diagonal entries are $A_1,\dots,A_r$, the remaining entries are zero.
\end{lem}
\begin{proof}
  *
\end{proof}

@@

Note from Equation~(\ref{eq:state-control}) that the control input $\mathbf{u}_k^a$ differs from the nominal control $\mathbf{u}_k$ in Equation~(\ref{eq:state-perf}). The first is an output of the algorithm. The second is a function that maps two consecutive control inputs to the difference in instantaneous energy consumption $\mathbf{u}_k:\mathbb{R}^{2\rho}\times\mathbb{Z}_{\geq 0}^{2\sigma}\rightarrow\mathbb{R}^{1+\rho}$.

@@

Specifically, the mission $\mathcal{M}$ contains a set of ordered lists with tasks (recall Definition~\ref{def:mission}). These tasks are implemented by software components (such as ROS nodes in a ROS based system) $\Psi(\mathbf{s}_i):=\langle\psi_1(s_{i,1}),\dots,\psi_\sigma(s_{i,\sigma})\rangle$. They input the desired and output the actual computations
\begin{equation}\label{eq:qos-def}
  \mathbf{s}_i:=\left\{\langle s_{i,1},\dots ,s_{i,\sigma}\rangle\mid \psi_j(s_{i,j})\in\mathbb{S}_{i}\,\forall j\in[\sigma]^+\right\},
\end{equation}
where $s_{i,j}:\mathbb{Z}_{\geq 0}\rightarrow\mathbb{Z}_{\geq 0}$ returns the $j$-th desired computation at stage $i$, $\mathbf{s}_i\in\mathbb{S}_i\subseteq\mathbb{Z}_{\geq 0}^\sigma$ the union of all QoS sets given $\mathbb{S}_i:=\bigcup_{j\in[\sigma]^+}{\mathbb{S}_{i,j}}$.

For instance, if the task $\psi_1$ is a CNN object detector, $s_{1,1}$ corresponds to the computation frames-per-second (fps) rate changing the detection frequency. The algorithm outputs the control input which contains the desired fps rate $s_{1,1}\in\mathbb{S}_{1,1}$. The task $\psi_1(s_{1,1})$ outputs the actual fps rate $s_{k,1}$ which might differ from the desired one (e.g., the CNN object detector might not be able to reach a high fps rate with the current computational resources).

Let us further define $g:\mathbb{Z}_{\geq 0}^{\sigma}\rightarrow\mathbb{R}_{\geq 0}$ as the instantaneous computational energy consumption value obtained using \powprof{}
\begin{equation}\label{eq:energy-comp}\begin{split}
  y_k^s:=g\left(\Psi(\mathbf{s}_i)\right)=g\left(\mathbf{s}_{k}\right).\\
\end{split}\end{equation}
Moreover, let $g(\{\emptyset\})$ be zero.

@@

Given $l$ stages $\mathcal{M}_i$ (TEE $\varphi_i$, tasks $\Psi$, computations $\mathbf{s}_i\in\mathbb{S}_i$, and adaptations $\mathbf{c}_i\in\mathbb{C}_i$ for all $i\in[l]$), the main purpose of the algorithm is to output a control input sequence $\mathbf{u}^a:=\{\mathbf{u}_0^a,\mathbf{u}_1^a,\dots\}$ in a valid mission.

@@

We stated earlier the output $y_k$--the instantaneous energy consumption--evolves in $\mathbb{R}_{\geq 0}$. This is generally untrue. Physical UAVs are bounded by strict energy budgets due to battery limitations.
